Semantic Novelty Detection with Graphical Models on Dora Architecture

### 2011-05-04 semantic-nd/article.tex
some changes

\keywords{novelty detection, semantic data, graphical models, room classification,
indoor environments, robotics.}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{itemize}
\item Explain context of indoor room categorization and importance of novelty detection.
\end{itemize}

@@

Dora\cite{dora} (CogX: Dora) was used as a base system where this novelty detection would be implemented.
As Dora moves through the environment its \emph{conceptual layer} builds a structural and probabilistic representation of the space
instantiated as a \emph{graphical model}.
That model connects the sensed properties together with the variables used to model the world.
And allows to perform queries on the probabilities of aspects of the environment.
For example: where is the robot most likely to find a cereal box\cite{exploiting}.

\subsection{Semantic Data}
\begin{itemize}
\item Explain what we mean by semantic data and how its captured by Dora
\end{itemize}

Dora has a set of machine learning tools that are used to sense the environment around it.
The output of those describe a certain high-level property or information that we refer as \emph{semantic data}.
The following types of probabilistic information are extracted by Dora and used on this paper:
\begin{description}
\item[Object detection] is performed by running several pre-trained object detectors on the visual input. The system becomes
aware of the presence of objects such as book, cereal box, computer, robot, stapler, toilet paper.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Novelty Detection}
\begin{itemize}
\item Explain 
\item Approach with a threshold function
\item Need of a normalizing factor for a threshold with dynamic functions
\item Usage of conditional density
\item Assuming of every output is equally likely
\item Using unlabelled data
\end{itemize}

>>>>

\keywords{novelty detection, semantic data, graphical models, room classification,
indoor environments, robotics, multi-modal.}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\begin{itemize}
\item Explain context of indoor room categorization and importance of novelty detection.
\item Introduce Dora and explain how it would benefit or use the semantic novel detection.
\end{itemize}

@@

Dora\cite{dora} (CogX: Dora) was used as a base system where this novelty detection would be implemented.
As Dora moves through the environment its \emph{conceptual layer} builds a structural and probabilistic representation of the space
instantiated as a \emph{graphical model}.
That model connects the sensed properties together with the variables used to model the world.
And allows to perform queries on the probabilities of aspects of the environment.
For example: where is most likely to find a cereal box\cite{exploiting}.

The ability to recognize novel categories on the modelled variables would allow to increase
reliability and allow the creation of self-extending behaviours.

\section{Related Work}
Although this paper deals with novelty detection on the robotics area, it does so using very
standard concepts and techniques such as semantic data and graphical models.
Those are often found on areas related with information retrieval.

Interesting examples are works in automatic image annotation using an hidden concept layer between visual features and text information\cite{multi-model-image-retrieval}.

\section{Dora Architecture Overview}
\begin{itemize}
\item Introduce Dora architecture aspects important for the paper
\item Explain what we mean by semantic data and how its captured by Dora
\item Introduce graphical models and how we use them (including notation on the paper)
\end{itemize}

Dora is composed by several layers, though for the sake of this paper only the conceptual layer is important.

\subsection{Semantic Data}
Dora has a set of machine learning tools that are used to sense the environment around it.
The output of those describe a certain high-level property or information that we refer as \emph{semantic data}.
The following types of probabilistic information are extracted by Dora and used on this paper:
\begin{description}
\item[Object detection] is performed by running several pre-trained object detectors on the visual input. The system becomes
aware of the presence of objects such as book, cereal box, computer, robot, stapler, toilet paper.

@@

The edges and factors between the several variables can be seen as a kind of filter to
the properties of the system.

Graphical models provide a generative model for a distributions of variables.
Those are mainly interesting since they allows to calculate the probability of any set of variables $x$

As they can be seen as a method to represent distributions of variables. From now on we shall use the notation $P_G(x)$ to denote the probability that the distribution described by a graphical model $G$ generates the set $x$ of variables.

\section{Novelty Detection}
\begin{itemize}
\item Explain 
\item Approach with a threshold function
\item Need of a normalizing factor for a threshold with dynamic functions
\item Usage of conditional density
\item Assuming of every output is equally likely
\item Using unlabelled data
\item What happens if the conditional and unconditional graphs differ only by the addition of a node?
\item How to implement that in the case a graph is a tree
\item Novelty detection on any variable of the graphical model!
\end{itemize}
